Robertson and Cao Project

### 2024-02-07 products/manuscript/Manuscript.qmd
Updated data set

---
title: "Robertson and Cao Project"
subtitle: "Teenage HPV Vaccination Coverage and Socioeconomic Factors"
author: Kelly Cao and Rachel Robertson
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../assets/dataanalysis-references.bib
csl: ../assets/apa.csl
---
# *Part 1*
## Data Set
  The data we chose to use is the National Immunization Survey (NIS) of 2022. It consists survey data that was collected to monitor vaccination coverage for teenagers age 13-17 in the United States of America. The surveys were conducted by the National Center for Immunization and Respiratory Diseases of the Center (NCIRD) for Disease Control and Prevention (CDC). The survey itself consists of two parts: (1) the household telephone (random digit dialing) survey answered by a parent or guardian and (2) a mailed survey for the vaccination provider, called “Immunization History Questionnaire.”
  The original data set is a fixed width file (FWF). In the data set, there are 43,738 rows and 672 columns, with each row representing an individual and the each column representing the answer to a question. The questions pertain to the teenager’s immunization history, demographics, and additional household-reported health information. 
We have truncated the data set to include 26-selected columns for the ease of observation. We have also filtered the data set to include one year, 2022, which is the most recent year of data that is published. The provided survey data will require cleaning, as there are signs of human errors present. 
## Opening the Data Set

First, we will load the readr package to open the file
```{r}
library(readr) #Loading the package to read the fwf
```

Next, we will define the column widths (beginning and end positions) and applying the column names to each position.
```{r}
column_position<- fwf_positions(
  start = c(1, 89, 287, 295, 298, 314, 330, 331, 334, 336, 343, 344, 345,
            347, 355, 359, 362, 472, 473, 495, 496, 497, 502, 504, 1296, 1297
  ),
  end = c(5, 92, 288, 295, 313, 329, 330, 332, 334, 336, 343, 344, 346,
          347, 356, 359, 362, 472, 473, 495, 496, 497, 502, 504, 1296, 1297
  ),
  col_names = c("SEQNUMT", "YEAR", "AGE", "EDUC1", "INCPORAR", "INCPORAR_I", 
                "INCPOV1", "INCQ298A", "LANGUAGE", "MOBIL_1", "RACEETHK", 
                "RACE_K", "RENT_OWN", "SEX", "STATE", "FACILITY", "WELLCHILD", 
                "P_U13HPV", "P_U13HPV3", "P_UTDHPV", "P_UTDHPV_15", 
                "P_UTDHPV_15INT", "P_UTDHPV2", "P_UTDHPV3", "INS_STAT2_I", 
                "INS_BREAK_I")
)

@@

Create the data frame "data" with read_fwf function. 
```{r}
data_frame <- read_fwf( "../products/NISTEENPUF22.DAT", 
                  col_positions = column_position, 
                  col_types = cols(.default = "c")
)
```

```{r}
#Read the statistics in the data file.
  summary(data_frame)
  str(data_frame)
```

The data will then be cleaned. As of now, the data set requires a library to translate the code.

The data will then be converted the data frame to a readable and downloadable .csv file for further analysis
```{r}
output_file<-"NIS_Teen_Data_2022.csv"
write.csv(data_frame, file = output_file, row.names = FALSE)
```

>>>>

---
title: "Robertson and Cao Project"
subtitle: "Part 1:Teenage HPV Vaccination Coverage and Socioeconomic Factors"
author: Kelly Cao and Rachel Robertson
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
---
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```
# Introduction
## Data Set
  The data we chose to use is the National Immunization Survey (NIS) of 2022. It consists survey data that was collected to monitor vaccination coverage for teenagers age 13-17 in the United States of America. The surveys were conducted by the National Center for Immunization and Respiratory Diseases of the Center (NCIRD) for Disease Control and Prevention (CDC). The survey itself consists of two parts: (1) the household telephone (random digit dialing) survey answered by a parent or guardian and (2) a mailed survey for the vaccination provider, called “Immunization History Questionnaire.”
  The original data set is a fixed width file (FWF). In the data set, there are 43,738 rows and 672 columns, with each row representing an individual and each column representing the answer to a question. The questions pertain to the teenager’s immunization history, demographics, and additional household-reported health information. 
We have truncated the data set to include 26-selected columns for the ease of observation. We have also filtered the data set to include one year, 2022, which is the most recent year of data that is published. The provided survey data will require cleaning, as there are signs of human errors present. 
## Opening the Data Set

First, we will load the readr package to open the file
```{r, echo=FALSE, message=FALSE}
library(readr) #Loading the package to read the fwf
library(here) #Loading the package to produce relative pathway
library(knitr) #Loading in the package for dynamic report generation
library(tidyverse)
library(readxl)
library(dplyr)
```

Next, we will define the column widths (beginning and end positions) and applying the column names to each position.
```{r, echo=FALSE, message=FALSE}
#Using the fwf_position() in the readr package to identify the columns widths
column_position<- fwf_positions(
  start = c(1, 89, 287, 295, 298, 314, 330, 331, 334, 336, 343, 344, 345,
            347, 355, 359, 362, 472, 473, 495, 496, 497, 502, 504, 1296, 1297
  ),
  end = c(5, 92, 288, 295, 313, 329, 330, 332, 334, 336, 343, 344, 346,
          347, 356, 359, 362, 472, 473, 495, 496, 497, 502, 504, 1296, 1297
  ),
  col_names = c("SEQNUMT", "YEAR", "AGE", "EDUC1", "INCPORAR", "INCPORAR_I", 
                "INCPOV1", "INCQ298A", "LANGUAGE", "MOBIL_1", "RACEETHK", 
                "RACE_K", "RENT_OWN", "SEX", "STATE", "FACILITY", "WELLCHILD", 
                "P_U13HPV", "P_U13HPV3", "P_UTDHPV", "P_UTDHPV_15", 
                "P_UTDHPV_15INT", "P_UTDHPV2", "P_UTDHPV3", "INS_STAT2_I", 
                "INS_BREAK_I")
)

@@

Create the data frame "data" with read_fwf function. 
```{r, echo=FALSE, message=FALSE}
data_file_path <- here("products", "manuscript", "NISTEENPUF22.DAT")

data_frame <- read_fwf(data_file_path, col_positions = column_position)

```{r}
EDUC1levels=c(1,2,3,4)
EDUC1labels=c("LESS THAN 12 YEARS", "12 YEARS", "MORE THAN 12 YEARS, NON-COLLEGE GRAD", "COLLEGE GRADUATE")

INCPOVlevels=c(1,2,3,4)
INCPOVlabels=c("ABOVE POVERTY > $75K", "ABOVE POVERTY <= $75K", "BELOW POVERTY", "UNKNOWN")

INCQ298Alevels=c(10,11,12,13,14,3,4,5,6,7,77,8,9,99)
INCQ298Alabels=c("$35001 - $40000", "$40001 - $50000", "$50001 - $60000", "$60001 - $75000", "$75001+", "$0 - $7500", "$7501 - $10000", "$10001 - $17500", "$17501 - $20000", "$20001 - $25000", "DON'T KNOW", "$25001 - $30000", "$30001 - $35000", "REFUSED")

LANGUAGElevels=c(1,2,3)
LANGUAGElabels=c("ENGLISH", "SPANISH", "OTHER")

MOBILlevels=c(1,2,77,98,99)
MOBILlabels=c("MOVED FROM DIFFERENT STATE", "DID NOT MOVE FROM DIFFERENT STATE", "DON'T KNOW", "MISSING IN ERROR", "REFUSED")

RACEETHKlevels=c(1,2,3,4)
RACEETHKlabels=c("HISPANIC", "NON-HISPANIC WHITE ONLY", "NON-HISPANIC BLACK ONLY", "NON-HISPANIC OTHER + MULTIPLE RACE")

RACE_Klevels=c(1,2,3)
RACE_Klabels=c("WHITE ONLY", "BLACK ONLY", "OTHER + MULTIPLE RACE")

RENTOWNlevels=c(1,2,3,77,99)
RENTOWNlabels=c("OWNED OR BEING BOUGHT", "RENTED", "OTHER ARRANGMENT", "DON'T KNOW", "REFUSED")

SEXlevels=c(1,2,77,98,99)
SEXlabels=c("MALE", "FEMALE", "DON'T KNOW", "MISSING IN ERROR", "REFUSED")

STATElevels=c(1,10,11,12,13,15,16,17,18,19,2,20,21,22,23,24,25,26,27,28,29,30,31,32,33,34,35,36,37,38,39,4,40,41,42,44,45,46,47,48,49,5,50,51,53,54,55,56,6,66,72,78,8,9)
STATElabels=c("ALABAMA", "DELAWARE", "DISTRICT OF COLUMBIA", "FLORIDA", "GEORGIA", "HAWAII", "IDAHO", "ILLINOIS", "INDIANA", "IOWA", "ALASKA", "KANSAS", "KENTUCKY", "LOUISIANA", "MAINE", "MARYLAND", "MASSACHUSETTS", "MICHIGAN", "MINNESOTA", "MISSISSIPPI", "MISSOURI", "MONTANA", "NEBRASKA", "NEVADA",
              "NEW HAMPSHIRE", "NEW JERSEY", "NEW MEXICO", "NEW YORK", "NORTH CAROLINA", "NORTH DAKOTA", "OHIO", "ARIZONA", "OKLAHOMA", "OREGON", "PENNSYLVANIA", "RHODE ISLAND", "SOUTH CAROLINA", "SOUTH DAKOTA", "TENNESSEE", "TEXAS", "UTAH", "ARKANSAS", "VERMONT", "VIRGINIA", "WASHINGTON", "WEST VIRGINIA",
              "WISCONSIN", "WYOMING", "CALIFORNIA", "GUAM", "PUERTO RICO", "U.S. VIRGIN ISLANDS", "COLORADO", "CONNECTICUT")

FACILITYlevels=c(1,2,3,4,5,6, ".")
FACILITYlabels=c("ALL PUBLIC FACILITIES", "ALL HOSPITAL FACILITIES", "ALL PRIVATE FACILITIES", "ALL STD/SCHOOL/TEEN CLINICS OR OTHER FACILITIES", "MIXED", "UNKNOWN", "Missing Data")

WELLCHILDlevels=c(1,2,3)
WELLCHILDlabels=c("YES", "NO", "DON'T KNOW")

UTDlevels=c(0,1, ".")
UTDlabels=c("NOT UTD", "UTD", "Missing Data")

INS_STAT2_Ilevels=c(1,2,3,4, ".")
INS_STAT2_Ilabels=c("PRIVATE INSURANCE ONLY", "ANY MEDICAID", "OTHER INSURANCE (CHIP, IHS, MILITARY, OR OTHER, ALONE OR IN COMB. WITH PRIVATE INSURANCE)", "UNINSURED", "MISSING Data")

INS_BREAK_Ilevels=c(1,2,3,4)
INS_BREAK_Ilabels=c("CURRENTLY INSURED BUT UNINSURED AT SOME POINT SINCE AGE 11", "CURRENTLY INSURED AND NEVER UNINSURED SINCE AGE 11", "CURRENTLY UNINSURED BUT INSURED AT SOME POINT SINCE AGE 11", "CURRENTLY UNINSURED AND NEVER INSURED SINCE AGE 11")

data<-mutate(data_frame,
             SEQNUMT = as.numeric(SEQNUMT), 
             YEAR = as.numeric(YEAR),
             AGE = as.numeric(AGE),
             EDUC1 = factor(EDUC1,levels = EDUC1levels, labels = EDUC1labels),
             INCPORAR_I = as.numeric(INCPORAR_I), 
             INCPOV1 = factor(INCPOV1, levels = INCPOVlevels, labels = INCPOVlabels),
             INCQ298A = factor(INCQ298A, levels = INCQ298Alevels, labels = INCQ298Alabels),             
             LANGUAGE = factor(LANGUAGE, levels = LANGUAGElevels, labels = LANGUAGElabels),
             MOBIL_1 = factor(MOBIL_1, levels = MOBILlevels, labels = MOBILlabels),
             RACEETHK = factor(RACEETHK, levels = RACEETHKlevels, labels = RACEETHKlabels),
             RACE_K = factor(RACE_K, levels = RACE_Klevels, labels = RACE_Klabels),
             RENT_OWN = factor(RENT_OWN, levels = RENTOWNlevels, labels = RENTOWNlabels), 
             SEX = factor(SEX, levels = SEXlevels, labels = SEXlabels),
             STATE = factor(STATE, levels = STATElevels, labels = STATElabels), 
             FACILITY = factor(FACILITY, levels = FACILITYlevels, labels = FACILITYlabels),
             WELLCHILD = factor(WELLCHILD, levels = WELLCHILDlevels, labels = WELLCHILDlabels),
             P_U13HPV = factor(P_U13HPV, levels = UTDlevels, labels = UTDlabels),
             P_U13HPV3 = factor(P_U13HPV3, levels = UTDlevels, labels = UTDlabels),
             P_UTDHPV = factor(P_UTDHPV, levels = UTDlevels, labels = UTDlabels),
             P_UTDHPV_15 = factor(P_UTDHPV_15, levels = UTDlevels, labels = UTDlabels),
             P_UTDHPV_15INT = factor(P_UTDHPV_15INT, levels = UTDlevels, labels = UTDlabels),
             P_UTDHPV2 = factor(P_UTDHPV2, levels = UTDlevels, labels = UTDlabels),
             P_UTDHPV3 = factor(P_UTDHPV3, levels = UTDlevels, labels = UTDlabels),
             INS_STAT2_I = factor(INS_STAT2_I, levels = INS_STAT2_Ilevels, labels = INS_STAT2_Ilabels),
             INS_BREAK_I = factor(INS_BREAK_I, levels = INS_BREAK_Ilevels, labels = INS_BREAK_Ilabels)
)

```{r, echo=FALSE, message=FALSE}
#Read the statistics in the data file.
  summary(data)
  str(data)
```

The data will then be cleaned.

The data will then be converted the data frame to a readable and downloadable .csv file for further analysis
```{r, echo=FALSE, message=FALSE}
output_file<-"NIS_Teen_Data_2022.csv"
write.csv(data, file = output_file, row.names = FALSE)
```
